---
title: "Tidymodels Classification Walkthrought"
subtitle: "using tidymodel"
author: "Giuliano Sposito"
output:
  md_document:
    variant: markdown_github
---

# Intro

This notebook tells the basic steps to ML pipeline using [`{tidymodels}`](https://www.tidymodels.org/) packages, this is a continuation of [Tidymodels Regression Walkthrough](./tidymodels_regression_walkthrough.md), but we'll do a classification in this case.

The traditional (without tunning) step for a ML pipeline are: 

1. Split the dataset between Training and Testing subsets ([`{rsample}`](https://rsample.tidymodels.org/))
1. Preprocessing and Feature Eng [`{recipes}`](https://recipes.tidymodels.org/)
1. Train a model ([`{parsnip}`](https://parsnip.tidymodels.org/)) using training dataset
1. Predict the outcome using the test dataset
1. Eval the model performance ([`{yardstick}`](https://yardstick.tidymodels.org/))

In this notebook we won't use the [`{workflow}`](https://workflows.tidymodels.org/) package, to understand the building blocks of 
[`{tidymodels}`](https://www.tidymodels.org/).

# (1) training and testing datasets

```{r classSplit, message=FALSE, warning=FALSE}

# dataset

# remotes::install_github("allisonhorst/palmerpenguins")
library(palmerpenguins)
library(skimr)
skim(penguins)


# tidymodel package to split datasets (tr/ts, CV,...)
library(rsample) 

# 1. train/test sets
penguins_split <- rsample::initial_split(penguins)
penguins_split

```

# (2) Preprocessing and Feature Eng

```{r classRecp, message=FALSE, warning=FALSE}

# tidymodel package to specify a sequence of transformation steps
library(recipes)

# transform/recipe
penguins_rec <- penguins_split %>%   # origim dataset
  training() %>%                     # training split
  recipe(species ~ ., data=.) %>%    # recipe template
  step_knnimpute(sex) %>%            # fill NA values of colum sex (factor)
  step_knnimpute(all_numeric()) %>%  # fill NA values of all numerical vars
  step_center(all_numeric()) %>%     # normalize to mean = 0
  step_scale(all_numeric()) %>%      # normalize to sd = 1
  step_mutate(year = factor(year, ordered=T)) %>%  # factor in to ordered
  step_dummy(c(island, sex)) %>%    # one hot encoding island and sex factors vars
  prep()                            # calculates recipe parameters

# datasets
penguins_test <- bake(penguins_rec, testing(penguins_split))
penguins_train <- juice(penguins_rec)

```


# (3) Training a model

```{r classTrain, message=FALSE, warning=FALSE}
# tidymodel package the uniforms the machine learnings algorithm interface
library(parsnip) # parsnip is the caret successor

# fit the model
penguins_model <- rand_forest(trees = 100, mode="classification") %>% 
  set_engine("ranger") %>% 
  fit(species ~ ., data=penguins_train)

penguins_model
```

# (4) Prediction

```{r classPred, message=FALSE, warning=FALSE}
# predict
predict(penguins_model, penguins_test)
```

# (5) Evaluate Model Performance

```{r classEval, message=FALSE, warning=FALSE}
# tidymodel package for measuring model performances
library(yardstick)

# eval
penguins_pred <- predict(penguins_model, penguins_test) %>%  # class outcome
  bind_cols(predict(penguins_model, penguins_test, type = "prob")) %>% # class probs
  bind_cols(penguins_test) %>% # true value
  relocate(species, everything())

head(penguins_pred)

# metrics
penguins_pred %>% 
  metrics(truth=species, estimate=.pred_class)

# checking AUC metrics
penguins_pred %>% 
  roc_auc(species, .pred_Adelie:.pred_Gentoo)

# plotting ROC curve
penguins_pred %>% 
  roc_curve(species, .pred_Adelie:.pred_Gentoo) %>% 
  tune::autoplot() # tune is a tidymodel package to find hyper-parameters

# checking gain
penguins_pred %>% 
  gain_curve(species, .pred_Adelie:.pred_Gentoo) %>% 
  tune::autoplot() # tune is a tidymodel package to find hyper-parameters

```
# full code

```{r fullCode, eval=FALSE}


# dataset

# remotes::install_github("allisonhorst/palmerpenguins")
library(palmerpenguins)
library(skimr)
skim(penguins)


# tidymodel package to split datasets (tr/ts, CV,...)
library(rsample) 

# 1. train/test sets
penguins_split <- rsample::initial_split(penguins)
penguins_split


# tidymodel package to specify a sequence of transformation steps
library(recipes)

# 2. transform/recipe
penguins_rec <- penguins_split %>%   # origim dataset
  training() %>%                     # training split
  recipe(species ~ ., data=.) %>%    # recipe template
  step_knnimpute(sex) %>%            # fill NA values of colum sex (factor)
  step_knnimpute(all_numeric()) %>%  # fill NA values of all numerical vars
  step_center(all_numeric()) %>%     # normalize to mean = 0
  step_scale(all_numeric()) %>%      # normalize to sd = 1
  step_mutate(year = factor(year, ordered=T)) %>%  # factor in to ordered
  step_dummy(c(island, sex)) %>%    # one hot encoding island and sex factors vars
  prep()                            # calculates recipe parameters

# datasets
penguins_test <- bake(penguins_rec, testing(penguins_split))
penguins_train <- juice(penguins_rec)

# tidymodel package the uniforms the machine learnings algorithm interface
library(parsnip) # parsnip is the caret successor

# 3. fit the model
penguins_model <- rand_forest(trees = 100, mode="classification") %>% 
  set_engine("ranger") %>% 
  fit(species ~ ., data=penguins_train)

penguins_model

# 4. predict
predict(penguins_model, penguins_test)

# tidymodel package for measuring model performances
library(yardstick)

# 5. eval
penguins_pred <- predict(penguins_model, penguins_test) %>%  # class outcome
  bind_cols(predict(penguins_model, penguins_test, type = "prob")) %>% # class probs
  bind_cols(penguins_test) %>% # true value
  relocate(species, everything())

head(penguins_pred)

# metrics
penguins_pred %>% 
  metrics(truth=species, estimate=.pred_class)

# checking AUC metrics
penguins_pred %>% 
  roc_auc(species, .pred_Adelie:.pred_Gentoo)

# plotting ROC curve
penguins_pred %>% 
  roc_curve(species, .pred_Adelie:.pred_Gentoo) %>% 
  tune::autoplot() # tune is a tidymodel package to find hyper-parameters

# checking gain
penguins_pred %>% 
  gain_curve(species, .pred_Adelie:.pred_Gentoo) %>% 
  tune::autoplot() # tune is a tidymodel package to find hyper-parameters

```

