---
title: "The Ames Housing Data EDA"
output: 
  md_document:
    toc: yes
editor_options: 
  chunk_output_type: console
---

```{r}
library(tidyverse) # eda and transformation
theme_set(theme_light())
```

# Dataset

In this chapter, we’ll introduce the Ames housing data set ([De Cock 2011](https://www.tmwr.org/ames#ref-ames)), which we will use in modeling examples throughout this book. Exploratory data analysis, like what we walk through in this chapter, is an important first step in building a reliable model. The data set contains information on 2,930 properties in Ames, Iowa, including columns related to:

. house characteristics (bedrooms, garage, fireplace, pool, porch, etc.)
. location (neighborhood)
. lot information (zoning, shape, size, etc.)
. ratings of condition and quality
. sale price

> Our modeling goal is to predict the sale price of a house based on other information we have, such as its characteristics and location.

```{r}
# load data from modeldata package
data(ames, package="modeldata")

dim(ames)
head(ames)
skimr::skim(ames)

```

# EDA

## Sale Price

```{r}
ames$Sale_Price |> 
  summary()

ames |> 
  ggplot(aes(x=Sale_Price)) +
  geom_histogram(bins=50, col="white", fill="darkblue") +
  scale_x_continuous(labels = scales::label_dollar()) +
  labs(title="Sales Price Distribution")
```

This plot shows us that the data are right-skewed; there are more inexpensive houses than expensive ones.

```{r}
ames |> 
  ggplot(aes(x=Sale_Price)) +
  geom_histogram(bins=50, col="white", fill="darkblue") +
  scale_x_continuous(labels = scales::label_dollar(), trans = scales::log10_trans()) +
  labs(title="Sales Price Distribution")

```

## Geographic Distribution

```{r}
ames |> 
  mutate( Sale_Price = log10(Sale_Price) ) |> 
  select(Neighborhood, Longitude, Latitude, Sale_Price) |> 
  ggplot() +
  geom_point(aes(x=Latitude, y=Longitude, color=Neighborhood, alpha=Sale_Price)) +
  theme_minimal() +
  theme(
    legend.position = "bottom"
  )
```

# Reference

All code and text came from Max Kuhn and Julia Silge`s book [Tidy Modeling with R](https://www.tmwr.org/ames).