---
title: "Ensembles of Models"
output: 
  md_document:
    toc: yes
editor_options: 
  chunk_output_type: console
---

A model ensemble, where the predictions of multiple single learners are aggregated to make one prediction, can produce a high-performance final model.

> Model stacking combines the predictions for multiple models of any type. For example, a logistic regression, classification tree, and support vector machine can be included in a stacking ensemble.

This chapter shows how to stack predictive models using the stacks package. We’ll re-use the results from [Chapter 15](./chapter15_workflowset.md) where multiple models were evaluated to predict the compressive strength of concrete mixtures.

```{r message=FALSE}
library(tidymodels)

# put rnotbook in the same workdir
knitr::opts_knit$set(root.dir = normalizePath(rprojroot::find_rstudio_root_file())) 

# recovering the training and test set
set.seed(1501)
concrete_split <- initial_split(concrete, strata = compressive_strength)
concrete_train <- training(concrete_split)
concrete_test  <- testing(concrete_split)

# recovering the tunned models
race_results <- readRDS("./chp15_race_results.rds")
race_results
```

The process of building a stacked ensemble is:

1. Assemble the training set of hold-out prediction (produced via resampling)
1. Create a model to blend these predictions.
1. For each member of the ensemble, fit ht emodel on the original training set.

> We’ll use the term candidate members to describe the possible model configurations (of all model types) that might be included in the stacking ensemble.

This means that a stacking model can include different types of models (e.g., trees and neural networks) as well as different configurations of the same model (e.g., trees with different depths).

## Creating the training set for stacking

The first step for building a stacked ensemble relies on the assessment set predictions from a resampling scheme with multiple splits. For each data point in the training set, stacking requires an out-of-sample prediction of some sort.

Back in Chapter 15, we used five repeats of 10-fold cross-validation to resample the data. This resampling scheme generates five assessment set predictions for each training set sample. 

> Simple validation sets can also be used with stacking since tidymodels considers this to be a single resample.

To start ensembling with the stacks package, create an empty data stack using the stacks() function and then add candidate models. Recall that we used workflow sets to fit a wide variety of models to these data. We’ll use the racing results:

```{r}
library(stacks)

concrete_stack <- 
  stacks() |> 
  add_candidates(race_results)

concrete_stack
```

Recall that racing methods are more efficient since they might not evaluate all configurations on all resamples. Stacking requires that all candidate members have the complete set of resamples. `add_candidates()` includes only the model configurations that have complete results.

## Blend the predictions

The training set predictions and the corresponding observed outcome data are used to create a meta-learning model where the assessment set predictions are the predictors of the observed outcome data. Meta-learning can be accomplished using any model. The most commonly used model is a regularized generalized linear model, which encompasses linear, logistic, and multinomial models.


Specifically, regularization via the lasso penalty (Tibshirani 1996), which uses shrinkage to pull points toward a central value, has several advantages:

+ Using the lasso penalty can remove candidates from ensemble.
+ The correlation between ensemble candidates tends to be very high, and regularization helps alleviate this issue.

```{r}
set.seed(2001)
ens <- blend_predictions(concrete_stack)

ens
```

This evaluates the meta-learning model over a predefined grid of lasso penalty values and uses an internal resampling method to determine the best value. 

```{r}
autoplot(ens)
```

The top panel shows the average number of candidate ensemble members retained by the meta-learning model. We can see that the number of members is fairly constant and, as it increases, the RMSE also increases.

The default range may not have served us well here. To evaluate the meta-learning model with larger penalties, let’s pass an additional option:

```{r}
set.seed(2002)
ens <- blend_predictions(concrete_stack,
                         penalty = 10^seq(-2, -0.5, length=20))

autoplot(ens)
```

The penalty value associated with the smallest RMSE was 0.051. Printing the object shows the details of the meta-learning model:

```{r}
ens
```

The regularized linear regression meta-learning model contained seven blending coefficients across four types of models. The autoplot() method can be used again to show the contributions of each model type, to produce.

```{r}
autoplot(ens, "weights") +
  geom_text(aes(x=weight+0.01, label=model), hjust=0) +
  theme_light() +
  theme(legend.position = "none") +
  lims(x=c(-0.01, 0.45))
```

## Fit the member models

The ensemble contains seven candidate members, and we now know how their predictions can be blended into a final prediction for the ensemble. However, these individual model fits have not yet been created. To be able to use the stacking model, seven additional model fits are required. These use the entire training set with the original predictors.

The `stacks` package has a function, `fit_members()`, that trains and returns these models:

```{r}
# remembering race_result is a tunned version of the model 
# training in the cross validation folds, not in the whole training dataset

library(rules) # necessary do the cubist model

ens <- fit_members(ens)
ens
```

## Test set results

Since the blending process used resampling, we can estimate that the ensemble with seven members had an estimated RMSE of 4.12. Recall from Chapter 15 that the best boosted tree had a test set RMSE of 3.41

```{r}
reg_metrics <- metric_set(rmse, rsq)

ens_test_pred <- 
  predict(ens, concrete_test) |> 
  bind_cols(concrete_test)

ens_test_pred |> 
  reg_metrics(compressive_strength, .pred)
```

This is moderately better than our best single model. It is fairly common for stacking to produce incremental benefits when compared to the best single model.

# Reference

All code and text came from Max Kuhn and Julia Silge`s book [Tidy Modeling with R](https://www.tmwr.org/ensembles).
